---
title: R語言期末競賽
author: G20
output:  html_document
---

<br><br>


##### Setup 


```{r}
packages = c(
  "dplyr","ggplot2","googleVis","devtools","magrittr","caTools","ROCR","caTools")
existing = as.character(installed.packages()[,1])
for(pkg in packages[!(packages %in% existing)]) install.packages(pkg)

if(!is.element("chorddiag", existing))
  devtools::install_github("mattflor/chorddiag")

```

##### Library
```{r echo=T, message=F, cache=F, warning=F}
rm(list=ls(all=T))
options(digits=4, scipen=12)
library(dplyr)
library(ggplot2)
library(caTools)
library(ROCR)
library(googleVis)
library(chorddiag)
```
<br><hr>







資料分割


<hr>

### Preparing The Predictors (X)
```{r echo=T, message=F, cache=F, warning=F}
Sys.setlocale("LC_TIME","C")
pacman::p_load(magrittr, readr, caTools, ggplot2, dplyr)
load("data/tf0.rdata")
```

##### The Demarcation Date
Remove data after the demarcation date
```{r}
feb01 = as.Date("2001-02-01")
Z = subset(Z0, date < feb01)    # 618212
```

##### Aggregate for the Transaction Records
```{r}
X = group_by(Z, tid) %>% summarise(
  date = first(date),  # 交易日期
  cust = first(cust),  # 顧客 ID
  age = first(age),    # 顧客 年齡級別
  area = first(area),  # 顧客 居住區別
  items = n(),                # 交易項目(總)數
  pieces = sum(qty),          # 產品(總)件數
  total = sum(price),         # 交易(總)金額
  gross = sum(price - cost)   # 毛利
  ) %>% data.frame  # 88387
```

```{r}
summary(X)
```

##### Check Quantile and Remove Outlier 
```{r}
sapply(X[,6:9], quantile, prob=c(.999, .9995, .9999))
```

```{r}
X = subset(X, items<=64 & pieces<=98 & total<=11260) # 88387 -> 88295
```

##### Aggregate for Customer Records
```{r}
d0 = max(X$date) + 1
A = X %>% mutate(
  days = as.integer(difftime(d0, date, units="days"))
  ) %>% 
  group_by(cust) %>% summarise(
    r = min(days),      # recency
    s = max(days),      # seniority
    f = n(),            # frquency
    m = mean(total),    # monetary
    rev = sum(total),   # total revenue contribution
    raw = sum(gross),   # total gross profit contribution
    age = age[1],       # age group
    area = area[1]     # area code
  ) %>% data.frame      # 28584
nrow(A)
```
<br><br><hr>

### Preparing the Target Variables (Y)

##### Aggregate Feb's Transaction by Customer
```{r}
feb = filter(X0, date>= feb01) %>% group_by(cust) %>% 
  summarise(amount = sum(total))  # 16899
```

##### The Target for Regression - `A$amount`
Simply a Left Joint
```{r}
A = merge(A, feb, by="cust", all.x=T)
```

##### The Target for Classification - `A$buy`
```{r}
A$buy = !is.na(A$amount)
```

##### Summary of the Dataset
```{r}
summary(A)
```

##### The Association of Categorial Predictors
```{r fig.height=3, fig.width=7.2}
tapply(A$buy, A$age, mean) %>% barplot
abline(h = mean(A$buy), col='red')
```

```{r fig.height=3, fig.width=7.2}
tapply(A$buy, A$area, mean) %>% barplot(las=2)
abline(h = mean(A$buy), col='red')
```

##### Contest Dataset
```{r}
X = subset(X, cust %in% A$cust & date < as.Date("2001-02-01"))
Z = subset(Z, cust %in% A$cust & date < as.Date("2001-02-01"))
set.seed(2018); spl = sample.split(A$buy, SplitRatio=0.7)
c(nrow(A), sum(spl), sum(!spl))
```

```{r fig.height=3, fig.width=7}
cbind(A, spl) %>% filter(buy) %>% 
  ggplot(aes(x=log(amount))) + geom_density(aes(fill=spl), alpha=0.5)
```


```{r}
A2 = subset(A, buy) %>% mutate_at(c("m","rev","amount"), log10)
n = nrow(A2)
set.seed(2018); spl2 = 1:n %in% sample(1:n, round(0.7*n))
c(nrow(A2), sum(spl2), sum(!spl2))
```

```{r fig.height=3, fig.width=7}
cbind(A2, spl2) %>% 
  ggplot(aes(x=amount)) + geom_density(aes(fill=spl2), alpha=0.5)
```


```{r}
save(Z, X, A, spl, spl2, file="data/tf2.rdata")
```

<br><br><hr>





建立模型

<hr>

### Loading & Preparing Data
```{r echo=T, message=F, cache=F, warning=F}
pacman::p_load(dplyr,ggplot2,caTools)
Sys.setlocale("LC_TIME","C")
load("data/tf2.rdata")
```

##### Spliting for Classification 
```{r}
TR = subset(A, spl)
TS = subset(A, !spl)
```
<br><hr>

### Classification Model
```{r}
glm1 = glm(buy ~ ., TR[,c(2:9, 11)], family=binomial()) 
summary(glm1)
pred =  predict(glm1, TS, type="response")
cm = table(actual = TS$buy, predict = pred > 0.5); cm
acc.ts = cm %>% {sum(diag(.))/sum(.)}; acc.ts          # 0.69998
colAUC(pred, TS$buy)                                   # 0.7556
```
<br><hr>

### Regression Model
```{r}
A2 = subset(A, A$buy) %>% mutate_at(c("m","rev","amount"), log10)
TR2 = subset(A2, spl2)
TS2 = subset(A2, !spl2)
```

```{r}
lm1 = lm(amount ~ ., TR2[,c(2:6,8:10)])
summary(lm1)
```

```{r}
r2.tr = summary(lm1)$r.sq
SST = sum((TS2$amount - mean(TR2$amount))^ 2)
SSE = sum((predict(lm1, TS2) -  TS2$amount)^2)
r2.ts = 1 - (SSE/SST)
c(r2.tr, r2.ts)
```
<br><hr>

### Prediction
```{r}
load("data/tf0.rdata")
d0 = max(X0$date) + 1
B = X0 %>% 
  filter(date >= as.Date("2000-12-01")) %>% 
  mutate(days = as.integer(difftime(d0, date, units="days"))) %>% 
  group_by(cust) %>% summarise(
    r = min(days),      # recency
    s = max(days),      # seniority
    f = n(),            # frquency
    m = mean(total),    # monetary
    rev = sum(total),   # total revenue contribution
    raw = sum(gross),   # total gross profit contribution
    age = age[1],       # age group
    area = area[1]     # area code
  ) %>% data.frame      # 28584
nrow(B)
```

```{r}
B$Buy = predict(glm1, B, type="response")
```

```{r}
B$Rev = predict(lm1, B)
```

```{r fig.height=2.4, fig.width=7.2}
par(mfrow=c(1,2), cex=0.8)
hist(B$Buy)
hist(log(B$Rev,10))
```

<br><br><hr><br><br>

```{r}
set.seed(111)
B = B[complete.cases(B),]
B$age = as.factor(B$age)
B$age = factor(B$age, order = T, levels = c("na", "a20", "a25", "a30", "a35", "a40", "a45", "a50", "a55", "a60", "a65"))
levels(B$age) = c(0, 20, 25, 30, 35, 40, 45, 50, 55, 60, 65)
B$age = as.character(B$age)
B$age = as.numeric(B$age)

B$grp = kmeans(scale(B[,c(2,4,5,7)]),10)$cluster
table(B$grp)  # 族群大小
B$grp = as.character(B$grp)
group_by(B, grp) %>% summarise(
  recent=mean(r), 
  freq=mean(f), 
  money=mean(m),
  age = mean(age),
  gross = mean(raw),
  size=n() ) %>% 
  mutate( revenue = size*money/1000 )  %>% 
  filter(size > 1)%>% 
  ggplot(aes(x=freq, y=money)) +
  geom_point(aes(size=revenue, col=recent),alpha=0.5) +
  scale_size(range=c(4,30)) +
  scale_color_gradient(low="green",high="red") +
  scale_x_log10() + scale_y_log10(limits=c(30,5000)) + 
  geom_text(aes(label = size ),size=3) +
  theme_bw() + guides(size=F) +
  labs(title="Customer Segements",
       subtitle="(bubble_size:revenue_contribution; text:group_size)",
       color="Recency") +
  xlab("Frequency (log)") + ylab("Average Transaction Amount (log)")

C = B %>% filter(grp==4)
summary(C)
```

```{r}
# my_customer = B %>% filter(grp==2)
# my_customer2 = left_join(my_customer, )
# my_customer2$week = as.factor(my_customer2$week)
# my_customer2$week = factor(my_customer2$week, order = T, levels = c('Monday', 'Tuesday', 'Wednesday', 'Thursday', 'Friday', 'Saturday', 'Sunday'))
# class(my_customer2$week)
```


#####行銷模擬，以下code需複製到Rscript才可執行
```{r}
# library(manipulate)
# manipulate({
#   do.call(rbind, lapply(seq(5,40,0.5), function(c){
#     p = m*plogis((10/a)*(c-b))
#     B %>% mutate(
#       PI = ifelse(B$Buy<=(1-p), p, 1-B$Buy) * B$Rev - c
#     ) %>%
#       group_by(grp) %>% summarise(
#         Cost = c,
#         Group.Sz = n(),
#         No.Target = sum(PI>0),
#         AvgROI = mean(PI[PI>0]),
#         TotalROI = sum(PI[PI>0])
#       ) } ) ) %>% 
#     ggplot(aes(x=Cost, y=TotalROI, col=grp)) +
#     geom_line(size=1.2) +
#     ggtitle("Cost Effeciency per Segment ")
# },
# m = slider(0.05, 0.25,  0.20, step=0.01),
# a = slider(  10,   30,    20, step=1),
# b = slider(   4,   20,    15, step=1)
# ) 
```

